---
title: "Taxonomic harmonization example"
output:
  html_document:
    df_print: paged
---

```{r laod libs and data, include=FALSE}
library(tidyverse)
library(rgnparser)
#library(rfishbase) needs to update R
library(lcvplants)

# act <- read_csv("/home/eb97ziwi/Proj/taxclean/data/biotime_actinopt.csv")[[1]]
plants <- read_csv("data/biotime_plants.csv")[[1]]
```

```{r dummy libs, eval=FALSE}
library(tidyverse)
library(rgnparser)
#library(rfishbase) needs to update R
library(lcvplants)

# act <- read_csv("/home/eb97ziwi/Proj/taxclean/data/biotime_actinopt.csv")[[1]]
plants <- unique(read_csv("data/biotime_plants.csv")[[1]])
```

## GNRparser

We start by pre-processing the plant list using `rgnrparser`, which removes abbreviations such as *sp.* and author names, resulting only in a canonical binomial name. This step is important as LCVP matches only names resolved at the species level (David said so, let's check maybe).

```{R gnrparser}
parsed <- gn_parse_tidy(plants) %>% 
  select(verbatim, canonicalfull) %>% 
  mutate(changed = ifelse(verbatim != canonicalfull, TRUE, FALSE),
         only_genus = modify(canonicalfull, function(x) {
           len <- str_split(x, " ", simplify = TRUE) %>% 
             length()
           ifelse (len == 1, TRUE, FALSE)
         }) %>% 
           as.logical())
# number of names changed
parsed %>% 
  group_by(changed) %>% 
  tally()
# example of changes
parsed %>% filter(changed)
```

## LCVP

We retain only species-level taxa and parse them to LCVP database using `lcvplants::LCVP()`.

```{r LCVP}
species <- parsed %>% 
  filter(!only_genus) %>% 
  pull(canonicalfull)
# backbone
plant_backbone <- tibble(bioTIME = species,
                         LCVP = sapply(species, function(x) tryCatch(
                           suppressMessages(lcvplants::LCVP(x)$LCVP_Accepted_Taxon),
                           error = function(e) NA)))
plant_backbone <- unnest(plant_backbone)
```
Let's have a look at what we got. First, how many species we find from the original species list.

```{r results}
plant_backbone <- plant_backbone %>% 
  mutate(matched = ifelse(LCVP == "" | is.na(LCVP), FALSE, TRUE))
plant_backbone %>% 
  group_by(matched) %>% 
  tally()
message(plant_backbone %>% filter(matched) %>% nrow(),
        "/", length(plants), " species found.")
```
We check how the new names compare to the original ones.

```{r comparison}
res <- tibble(bioTIME = plants) %>% 
  left_join(parsed %>% 
              transmute(bioTIME = verbatim,
                        parsed = canonicalfull)) %>% 
  left_join(plant_backbone %>% 
              filter(matched) %>% 
              transmute(parsed = bioTIME,
                        LCVP))
res
```

We can keep only the canonical binomial name using `rgneparser`, just to compare programmatically the difference.

```{r remove authors}
res <- res %>% mutate(LCVP_parsed = gn_parse_tidy(res$LCVP)$canonicalfull) %>% 
  mutate(changed = ifelse(bioTIME == LCVP_parsed, FALSE, TRUE) %>% 
           replace_na(FALSE))
res %>% 
  group_by(changed) %>% 
  tally()
res %>% filter(changed)
```

